---
title: "Homework 3"
author: "Iman Malik"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      eval = FALSE, 
                      warning = FALSE,
                      message = FALSE,
                      fig.align = "center")
```

```{r}
# Datasets and Library
library(ggplot2)
library(tidyverse)
library(mosaic)

gas_prices <- read.csv("gasprices.csv")
sclass <- read.csv("sclass.csv")
nbc_pilot <- read.csv("nbc_pilotsurvey.csv")
ebay <- read.csv("ebay.csv")
```

UT EID: im22856

GitHub repo: <https://github.com/imanmalik16/SDS-315.git>

## Problem 1

### a)

```{r}
boot_competition = do(10000) * diffmean(Price ~ Competitors, data = resample(gas_prices))
confint(boot_competition, level = 0.95)
```

Theory A suggests that gas stations charge more if they lack direct competition in sight. We can say with 95% confidence that the difference in mean gas prices between gas stations with and without direct competition in sight is between -6 cents and 1 cent, rounded to the nearest cent. This difference is extremely small and trivial in the context of prices, so there is no practical significance. Additionally, it is not statistically significant at the 5% level because the 95% confidence interval for this difference in means does contain zero. Due to the lack of statistical significance, there is not convincing evidence to support this theory.

### b)

```{r}
gas_prices_b <- gas_prices |>
  mutate(Wealth = ifelse(Income >= median(Income), yes = TRUE, no = FALSE))
boot_income = do(10000) * diffmean(Price ~ Wealth, data = resample(gas_prices_b))
confint(boot_income, level = 0.95)
```

Theory B suggests that the richer the area, the higher the gas prices. We can say with 95% confidence that the difference in mean gas prices between gas stations in wealthy and non-wealthy areas is between 0 cents and 6 cents, rounded to the nearest cent (the lower bound rounded down to 0, but 0 is not included in this interval). We categorized wealthy areas as areas with household income greater than or equal to the median income of the entire dataset. This difference is extremely small and trivial in the context of prices, with gas stations in wealthier areas charging a mere few cents more for gas, so there is no practical significance. However, it is statistically significant at the 5% level because the 95% confidence interval for this difference in means does not contain zero. Due to this statistical significance, there is convincing evidence to support this theory.

### c)

```{r}
boot_stoplight = do(10000) * diffmean(Price ~ Stoplight, data = resample(gas_prices))
confint(boot_stoplight, level = 0.95)
```

Theory C suggests that gas stations at stoplights charge more. We can say with 95% confidence that the difference in mean gas prices between gas stations with and without a stoplight is between -4 cents and 3 cents, rounded to the nearest cent. This difference is extremely small and trivial in the context of prices so there is no practical significance. Additionally, it is not statistically significant at the 5% level because the 95% confidence interval for this difference in means does contain zero. Due to the lack of statistical significance, there is not convincing evidence to support this theory.

### d)

```{r}
boot_highway = do(10000) * diffmean(Price ~ Highway, data = resample(gas_prices))
confint(boot_highway, level = 0.95)
```

Theory D suggests that gas stations with direct highway access charge more. We can say with 95% confidence that the difference in mean gas prices between gas stations with and without direct access to a highway is between 1 cent and 8 cents, rounded to the nearest cent. This difference is extremely small and trivial in the context of prices, with gas stations with highway access charging a mere few cents more for gas, so there is no practical significance. However, it is statistically significant at the 5% level because the 95% confidence interval for this difference in means does not contain zero. Due to this statistical significance, there is convincing evidence to support this theory.

### e)

```{r}
gas_prices_e <- gas_prices |>
  mutate(isShell = str_detect(Name, pattern = "Shell"))

boot_shell = do(10000) * diffmean(Price ~ isShell, data = resample(gas_prices_e))
confint(boot_shell, level = 0.95)
```

Theory E suggests that Shell charges more than all other non-Shell brands. We can say with 95% confidence that the difference in mean gas prices between Shell gas stations and gas stations of other brands is between -1 cent and 7 cents, rounded to the nearest cent. This difference is extremely small and trivial in the context of prices so there is no practical significance. Additionally, it is not statistically significant at the 5% level because the 95% confidence interval for this difference in means does contain zero. Due to the lack of statistical significance, there is not convincing evidence to support this theory.

## Problem 2

### a)

```{r}
sclass_a <- sclass |>
  filter(year == 2011, trim == "63 AMG")

boot_mileage = do(10000) * mean(~mileage, data = resample(sclass_a))
confint(boot_mileage, level = 0.95)
```

We can say with 95% confidence that the average mileage of 2011 S-Class 63 AMGs on the used-car market is somewhere between 26230.93 miles and 31846.14 miles.

### b)

```{r}
sclass_b <- sclass |>
  filter(year == 2014, trim == "550") |>
  mutate(isBlack = str_detect(color, pattern = "Black"))

boot_proportion = do(10000) * prop(~isBlack, data = resample(sclass_b))
confint(boot_proportion, level = 0.95)
```

We can say with 95% confidence that the proportion of all 2014 S-Class 550s that were painted black on the used-car market is somewhere between 41.64% and 45.28%..

## Problem 3

### a)

```{r}
nbc_pilot_a <- nbc_pilot |>
  filter(Show == "Living with Ed" | Show == "My Name is Earl")

boot_happy = do(10000) * diffmean(Q1_Happy ~ Show, data = resample(nbc_pilot_a))
confint(boot_happy, level = 0.95)
```

1)  Is there evidence that one show consistently produces a higher mean Q1_Happy response among viewers?
2)  I created 10,000 bootstrap samples from the filtered nbc_pilot dataset, and computed the difference in the means of the Q1_Happy variable between "Living with Ed" and "My Name is Earl" for each bootstrap sample. I then stored this in the boot_happy variable, which I used to construct a 95% confidence interval.
3)  I can say with 95% confidence that the difference in average Q1_Happy responses between the two shows is between -0.400 and 0.099.
4)  The difference in average Q1_Happy responses between the two shows is not statistically significant because the 95% confidence interval for this difference does contain zero. Therefore, there is not convincing evidence that one show consistently produces a higher mean Q1_Happy response among viewers.

### b)

```{r}
nbc_pilot_b <- nbc_pilot |>
  filter(Show == "The Biggest Loser" | Show == "The Apprentice: Los Angeles")

boot_annoyed = do(10000) * diffmean(Q1_Annoyed ~ Show, data = resample(nbc_pilot_b))
confint(boot_annoyed, level = 0.95)
```

1)  Is there evidence that one show consistently produces a higher mean Q1_Annoyed response among viewers?
2)  I created 10,000 bootstrap samples from the filtered nbc_pilot dataset, and computed the difference in the means of the Q1_Annoyed variable between "The Biggest Loser" and "The Apprentice: Los Angeles" for each bootstrap sample. I then stored this in the boot_annoyed variable, which I used to construct a 95% confidence interval.
3)  I can say with 95% confidence that the difference in average Q1_Annoyed responses between the two shows is between -0.521 and -0.016.
4)  The difference in average Q1_Annoyed responses between the two shows is statistically significant because the 95% confidence interval for this difference does not contain zero. Therefore, there is convincing evidence that "The Apprentice: Los Angeles" produces a higher mean Q1_Annoyed response among viewers than "The Biggest Loser".

### c)

```{r}
nbc_pilot_c <- nbc_pilot |>
  filter(Show == "Dancing with the Stars") |>
  mutate(isConfusing = Q2_Confusing >= 4)

boot_confusing = do(10000) * prop(~isConfusing, data = resample(nbc_pilot_c))
confint(boot_confusing, level = 0.95)
```

1)  Based on this sample of respondents, what proportion of American TV watchers would we expect to give a response of 4 or greater to the “Q2_Confusing” question?
2)  I first created a new logical variable using mutate() to indicate whether a show was rated to be confusing or not, and I filtered it to only include the show "Dancing with the Stars". I then created 10,000 bootstrap samples from the filtered nbc_pilot dataset, and computed the proportion of watchers who rated the show to be annoying for each bootstrap sample. I then stored this in the boot_confusing variable, which I used to construct a 95% confidence interval.
3)  We can say with 95% confidence that the proportion of Q1_Confused responses that agree "Dancing with the Stars" is a confusing show is between 3.87% and 12.15%.
4)  The proportion of American TV watchers we would expect to give a response of 4 or greater (indicating confusion) on the Q2_Confusing question is rather low. Therefore, only a small percentage of viewers find the show confusing.

## Problem 4

```{r}
ebay <- ebay |>
  mutate(revenue_ratio = rev_after / rev_before)

boot_diff <- do(10000) * diffmean(revenue_ratio ~ adwords_pause, data = resample(ebay))
confint(boot_diff, level = 0.95)
```

1)  Does the extra traffic brought to our site from paid search results--above and beyond whatwe’d see if we “went organic”--justify the cost of the ads themselves?
2)  I first created a new numeric variable using mutate() that shows the ratio of revenue after to revenue before for each DMA. I then created 10,000 bootstrap samples from the altered ebay dataset, and computed the difference in means of the revenue ratio between the control group and the treatment group. I then stored this in the boot_diff variable, which I used to construct a 95% confidence interval.
3)  I can say with 95% confidence that the difference in means of the revenue ratio between the treatment group and the control group is between -0.09 and -0.01.
4)  This difference is statistically significant at the 5% level because the 95% confidence interval does not contain zero. Therefore, there is convincing evidence that the control group produces a higher mean revenue ratio than the treatment group. In the context of the question, the cost of paying for search results is not justified because "going organic" resulted in a higher revenue ratio than paying to bring in extra traffic to the site.
